---
title: "Generating Hypsography"
author: "Joseph Stachelek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Hypsography}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The General Lake Model (GLM) requires information on the depth-area relationship for a given lake. In GLM this is referred to as hypsography (rather than bathymetry) and has the following description in the `glmtools` documentation for `get_hypsography`:  

<br>

```{r echo=FALSE, results='hide', message=FALSE}
library(glmtools)
```

```{r echo=FALSE}
library(printr)
```

```{r tidy=FALSE, printr.help.sections=c('description'), comment='', echo=FALSE}
?get_hypsography
```

----

## Example

Let's start by investigating (reverse-engineering) the hypsography provided for Sparkling Lake (Wisconsin) in the `run_example_sim` function.

```{r }
sim_folder <- run_example_sim(verbose = FALSE)
nml_file <- file.path(sim_folder, 'glm2.nml')
get_hypsography(file = nml_file)
```

```{r echo=FALSE}
hp <- get_hypsography(file = nml_file)[,2:1]
plot(hp, ylim = rev(range(hp[,2])))
```

### Verify max depth

```{r }
# Get max depth
tail(hp, 1)
```

This value matches the max depth reported by the [Wisconsin DNR](http://dnr.wi.gov/lakes/lakepages/LakeDetail.aspx?wbic=1881900&page=facts).

### Verify dimensions

```{r }
nml <- read_nml(nml_file)
get_nml_value(nml, "bsn_len")
get_nml_value(nml, "bsn_wid")
```

```{r }
library(lakemorpho)
library(nhdR)


```

## Generating from scratch