---
title: "Generating Hypsography"
author: "Joseph Stachelek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Hypsography}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The General Lake Model (GLM) requires information on the depth-area relationship for a given lake. In GLM this is referred to as hypsography (rather than bathymetry) and has the following description in the `glmtools` documentation for `get_hypsography`:  

<br>

```{r echo=FALSE, results='hide', message=FALSE}
library(glmtools)
```

```{r echo=FALSE}
library(printr)
```

```{r tidy=FALSE, printr.help.sections=c('description'), comment='', echo=FALSE}
?get_hypsography
```

----

## Example

Let's start by investigating (reverse-engineering) the hypsography provided for Sparkling Lake (Wisconsin) in the `run_example_sim` function.

```{r eval = FALSE}
sim_folder <- run_example_sim("inst/extdata", verbose = FALSE)
nml_file <- file.path("inst/extdata", "glm2.nml")
```

```{r echo=FALSE}
nml_file <- file.path("../../inst/extdata", "example_sim_sparkling.nml")
```

```{r }
nml <- read_nml(nml_file)
hp <- get_hypsography(file = nml_file)[,2:1]
```

```{r echo=FALSE}
# plot(hp, ylim = rev(range(hp[,2])))
```

### Calculate volume from hypsographic profile

```{r }
hp_diff <- cbind(c(diff(hp$depths), 0), hp$areas)

calc_layer_vol <- function(hp_diff, i){
  top_area  <- hp_diff[i, 1] 
  bot_area  <- hp_diff[i + 1, 2]
  # volume of a layer approximated as a truncated cone
  (hp_diff[i, 1] / 3) * (top_area + bot_area + sqrt(top_area * bot_area))
}

vol_layers <- sapply(seq_len(nrow(hp_diff) - 1), 
                            function(i) calc_layer_vol(hp_diff, i))
(vol_estimated <- sum(vol_layers))
```
  
### Approximate hypsographic profile from max depth and surface area

```{r }
(nlayers <- nrow(hp))
(max_depth <- tail(hp$depths, 1))
(surface_area <- head(hp$areas, 1))

res <- suppressWarnings(
  glmutils::generate_hypsography(nlayers, max_depth, surface_area))

plot(res[,c("area", "depth_linear")],
     ylim = c(max(res$depth_linear), min(res$depth_linear)), 
     ylab = "depth")
points(res[,c("area", "depth_ellipsoid")], col = "red", pch = 19)
legend("topleft", legend = c("linear", "ellipsoid"), col = c("black", "red"), 
        pch = c(21, 19))
```


### Verify max depth

```{r }
# Get max depth
tail(hp, 1)
```

This value matches the max depth reported by the [Wisconsin DNR](http://dnr.wi.gov/lakes/lakepages/LakeDetail.aspx?wbic=1881900&page=facts).

### Verify dimensions

```{r verify dimensions}
get_nml_value(nml, "bsn_len")
get_nml_value(nml, "bsn_wid")
```

```{r cache=TRUE}
library(nhdR)
# Get NHD Waterbody Polygons
# nhd_plus_get(vpu = 7, "NHDSnapshot")
dt <- nhd_plus_load(vpu = 7, "NHDSnapshot", "NHDWaterbody")
spark_outline <- dt[grep("Sparkling", dt$GNIS_NAME),]
plot(spark_outline$geometry, main = "Sparkling Lake")

suppressPackageStartupMessages(library(lakemorpho))
library(raster)
# Get length and width
spark_outline_sp <- as(spark_outline, "Spatial")
r <- rasterize(spark_outline_sp, raster(spark_outline_sp))

spark_outline_sp <- spTransform(spark_outline_sp, CRS("+proj=utm +zone=10 +datum=WGS84"))
r <- projectRaster(r, crs = "+proj=utm +zone=10 +datum=WGS84")

spark_lm <- lakeSurroundTopo(spark_outline_sp, r)

calcLakeMetrics(spark_lm, bearing = 45, pointDens = 250)
```
